Species richness and network structure jointly drive total biomass and its temporal stability in fish communities

### 2020-12-07 manuscript/bef_stability/manuscript.Rmd
Add new references: fish biomass, stability

Both biodiversity-ecosystem functioning (BEF) and foodweb complexity-stability
relationships have received considerable attention during the last decades.
Still, the BEF relationship remains rarely investigated in the context of food
webs, especially in empirical studies explicitly considering foodweb structures.
Foodweb stability-complexity relationship has been mainly the focus of
theoretical studies with very little empirical testing. To fill this gap, we linked
species richness, foodweb structure, annual biomass and its temporal
stability using a standardized monitoring of fish communities in French streams
running from 1995 to 2018. We inferred trophic interactions from
body size recorded and diet literature. We reveal that species richness and average
trophic level had a positive relationship with annual biomass of fish
communities while they had respectively a negative and positive effect on its
temporal stability. We also found that the effects of stream size and
temperature on annual biomass and stability are mediated by their effects on
diversity and foodweb structure. Overall, our study opens fruitful perspectives
on the link among species richness, network structure, community biomass and its
temporal stability.

@@

Overall, the evidence suggests that both species richness and network structure
should modulate total biomass and stability of communities. The effects of
species richness and network structure, such as connectance and average trophic
level, are potentially conflicting or synergistic in such a way that they need
to be addressed jointly. Despite the long-standing interest for the links
between foodweb complexity and stability in ecology, we still have very few
evidence of the relationship between network structure and the temporal
stability of multitrophic communities in natural settings.
Although the relationships between diversity, total biomass and its stability
found in experimental and in natural settings can be similar
[e.g., @wu_relationship_2015; @li_relationship_2018], negative or non
significant effects of species richness on total biomass have been found in some
natural communities such as forests [@szwagrzyk_above-ground_2007;
@jacob_productivity_2010; @seidel_relationship_2013]. Some authors suggest that
the lack of resource complementarity or environmental harshness might be
responsible for negligible, or negative effects of species richness
[@loreau_biodiversity_2001; @he_global_2013; @wu_relationship_2015].

@@

When analyzing data from natural communities, the relationships among species
richness, total biomass and species richness might also arise from effects of
the environment on both diversity and ecosystem functions
[@loreau_biodiversity_2001; @he_global_2013;
@grace_integrative_2016]. Environment affects organism performance directly, and
thereby, species abundances, total biomass [@grace_integrative_2016] and temporal
stability [@de_boeck_patterns_2017; @hansen_climate_2013; @bluthgen_land_2016;
@oliver_heterogeneous_2010; @franssen_annual_2011; @friedpetersen_drivers_2020].
For example, larger ecosystem might increase the number of available
resources and thus total biomass [@post_ecosystem_2000; @doi_resource_2009].
@li_relationship_2018 have notably shown that climate moisture has a positive
direct effect on total biomass, but also indirect through a positive effect on
species richness. Accounting for environment also allow for an overall
comparison of the strength of the effects of environment and community structure
on total biomass and its stability.

@@

We found that species richness increased total biomass as found in salt marshes,
grasslands and forests [@hector_plant_1999; @tilman_biodiversity_2014;
@wu_relationship_2015; @li_relationship_2018]. The positive effect of species
richness on total biomass might be due to resource use complementarity
[@carey_determining_2011], which is held by niche differentiation
[@loreau_biodiversity_1998; @loreau_partitioning_2001] or to the higher
probability of getting high biomass species with higher species richness
[@loreau_biodiversity_2001].

@@

Consistent with research showing that foodwebs in wider streams support more
species and biomass [@allan_stream_2007], we found that the PCA axis related to
stream size had a positive effect on total
biomass through a positive effect on mean trophic level. Greater ecosystem size
might provide a greater diversity of resources through habitat heterogeneity,
and so higher species richness and total biomass by surface unit, through
resource complementarity. Greater diversity of resources can also maintain
more biomass at high trophic level [@post_ecosystem_2000; @doi_resource_2009].
Furthermore, the positive effect of the PCA axis related to temperature on total
biomass was of the magnitude of effects of network structure and species
richness, in line with the mounting body of work reporting important effects of
temperature on the structure and biomass of aquatic system [e.g.
@dossena_warming_2012; @edeline_ecological_2013; @emmrich_geographical_2014;
@hattab_forecasting_2016]. Contrasting with a previous study in experimental
plant communities [@grace_integrative_2016] and in fish communities
[@yvondurocher_warming_2011; @dossena_warming_2012], we found that environmental
drivers overall affect total biomass indirectly rather than directly.

>>>>

Both biodiversity-ecosystem functioning (BEF) and foodweb complexity-stability
relationships have received considerable attention during the last decades.
Still, the BEF relationship remains rarely investigated in the context of food
webs, especially in empirical studies explicitly considering foodweb structures.
Foodweb stability-complexity relationship has been mainly the focus of
theoretical studies with very little empirical testing. To fill this gap, we linked
species richness, foodweb structure, annual biomass and its temporal
stability using a standardized monitoring of fish communities in French streams
running from 1995 to 2018. We inferred trophic interactions from
individual body size and diet literature. We reveal that species richness and average
trophic level had a positive relationship with annual biomass of fish
communities while they had respectively a negative and positive effect on its
temporal stability. We also found that the effects of stream size and
temperature on annual biomass and stability are mediated by their effects on
diversity and foodweb structure. Overall, our study opens fruitful perspectives
on the link among species richness, network structure, community biomass and its
temporal stability.

@@

Overall, the evidence suggests that both species richness and network structure
should modulate total biomass and stability of communities. The effects of
species richness and network structure, such as connectance and average trophic
level, are potentially conflicting or synergistic in such a way that they need
to be addressed jointly. Despite the long-standing interest for the links
between foodweb complexity and stability in ecology, we still have very few
evidence of the relationship between network structure and the temporal
stability of multitrophic communities in natural settings.
Although the relationships between diversity, total biomass and its stability
found in experimental and in natural settings can be similar
[e.g., @wu_relationship_2015; @li_relationship_2018], negative or non
significant effects of species richness on total biomass have been found in some
natural communities such as forests [@szwagrzyk_above-ground_2007;
@jacob_productivity_2010; @seidel_relationship_2013] and marine fishes [@maureaud_biodiversityecosystem_2019]. Some authors suggest that
the lack of resource complementarity, environmental harshness or past human
pressures might be responsible for negligible, or negative effects of species
richness [@loreau_biodiversity_2001; @he_global_2013; @wu_relationship_2015;
@maureaud_biodiversityecosystem_2019].

@@

When analyzing data from natural communities, the relationships among species
richness, total biomass and species richness might also arise from effects of
the environment on both diversity and ecosystem functions
[@loreau_biodiversity_2001; @he_global_2013;
@grace_integrative_2016]. Environment affects organism performance directly, and
thereby, species abundances, total biomass [@grace_integrative_2016; @woods_testing_2020] and temporal
stability [@de_boeck_patterns_2017; @hansen_climate_2013; @bluthgen_land_2016;
@oliver_heterogeneous_2010; @franssen_annual_2011; @friedpetersen_drivers_2020].
For example, larger ecosystem might increase the number of available
resources and thus total biomass [@post_ecosystem_2000; @doi_resource_2009].
@li_relationship_2018 have notably shown that climate moisture has a positive
direct effect on total biomass, but also indirect through a positive effect on
species richness. Accounting for environment also allow for an overall
comparison of the strength of the effects of environment and community structure
on total biomass and its stability.

@@

We found that species richness increased total biomass as found in salt marshes,
grasslands and forests [@hector_plant_1999; @tilman_biodiversity_2014;
@wu_relationship_2015; @li_relationship_2018] as well as in streams
[@woods_testing_2020]. The positive effect of species
richness on total biomass might be due to resource use complementarity
[@carey_determining_2011], which is held by niche differentiation
[@loreau_biodiversity_1998; @loreau_partitioning_2001] or to the higher
probability of getting high biomass species with higher species richness
[@loreau_biodiversity_2001].

@@

Consistent with research showing that foodwebs in wider streams support more
species and biomass [@allan_stream_2007], we found that the PCA axis related to
stream size had a positive effect on total
biomass through a positive effect on mean trophic level. Greater ecosystem size
might provide a greater diversity of resources through habitat heterogeneity,
and so higher species richness and total biomass by surface unit, through
resource complementarity. Greater diversity of resources can also maintain
more biomass at high trophic level [@post_ecosystem_2000; @doi_resource_2009].
Furthermore, the positive effect of the PCA axis related to temperature on total
biomass was of the magnitude of effects of network structure and species
richness, in line with the mounting body of work reporting important effects of
temperature on the structure and biomass of aquatic system [e.g.
@dossena_warming_2012; @edeline_ecological_2013; @emmrich_geographical_2014;
@hattab_forecasting_2016; @woods_testing_2020]. Contrasting with previous studies in experimental
plant communities [@grace_integrative_2016], aquatic mesocosms [@yvondurocher_warming_2011; @dossena_warming_2012] and marine fish communities [@maureaud_biodiversityecosystem_2019]
, we found that environmental drivers overall affect total biomass indirectly
rather than directly [@woods_testing_2020].
